# Homework 1: Basic Data Wrangling

Please use the pipe operator to represent composite tasks. 

1. Load the dataset `chicago.rds` from moodle using the readRDS() function of R into a dataframe called `chicago`. Don't display the code to laod the file in the knitted document. The dataset contains air pollution data from Chicago from a period Jan 1987 till end of 2005. To learn more about these measures visit http://www.airnow.gov/index.cfm?action=aqibasics.particle   
```{r,include=FALSE}
library (dplyr)
library(ggplot2)
library(zoo)
```

```{r, echo = FALSE}
chicago <- readRDS('C:/Users/Pranathi/Desktop/MSBA/Fall/Exploratory data analytics/Assignment 1/chicago.rds')
```

2. Make sure the dataset is a dplyr object from here on. Perform basic data cleansing activities that deal with outliers, data entry errors and missing values if any.

```{r}
chicago_table <- tbl_df(chicago)
summary(chicago_table)
```

Missing value count: 
tmpd - 1
dptp - 2
pm25tmean2 - 4447
pm10tmean2 - 242
Total number of Records - 6940

Missing Value Treatment:
1. For Temperature and dew point temperature, missing values are replaced with mean of previous and next day's temperature/dew point temperature as we can safely assume that the temperature doesn't change drastically within a day's time
2. For PM2.5, linear regression is used to predict the missing values
Regression has been used as other techniques are not accurate given that 64% of the records are missing values. Below are the reasons why other techniques cannot be used
a) mean of nearby values - most of the records will not have nearby values 
b) omitting -more than 50% data loss 
c) Replacing with column mean - Loss of variation in data as ~60% of the data will be replaced with a single value

```{r,warning=FALSE}
chicago_table <- arrange(chicago_table,date)

# Replacing missing values in Temperature and dew point temperature with mean of nearby values
chicago_table$tmpd <-  (na.locf(chicago_table$tmpd) + rev(na.locf(rev(chicago_table$tmpd))))/2
chicago_table$dptp <- (na.locf(chicago_table$dptp) + rev(na.locf(rev(chicago_table$dptp))))/2

#Outlier Inspection
ggplot(chicago_table) + geom_boxplot(aes(x=city,y=pm25tmean2))
ggplot(chicago_table) + geom_boxplot(aes(x=city,y=pm10tmean2))
ggplot(chicago_table) + geom_boxplot(aes(x=city,y=o3tmean2))
ggplot(chicago_table) + geom_boxplot(aes(x=city,y=no2tmean2))

#Replace outliers with NA to create a training dataset
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 3 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

chicago_outliers <- chicago_table
chicago_outliers$tmpd <- remove_outliers(chicago_outliers$tmpd)
chicago_outliers$dptp <- remove_outliers(chicago_outliers$dptp)
chicago_outliers$pm25tmean2 <- remove_outliers(chicago_outliers$pm25tmean2)
chicago_outliers$pm10tmean2 <- remove_outliers(chicago_outliers$pm10tmean2)
chicago_outliers$o3tmean2 <- remove_outliers(chicago_outliers$o3tmean2)
chicago_outliers$no2tmean2 <- remove_outliers(chicago_outliers$no2tmean2)

#Train data set - remove missing values and outliers
train_data <- na.omit(chicago_outliers)

#model to predict pm10tmean2 
m1_pm10 <-lm(formula = pm10tmean2 ~ tmpd+dptp+o3tmean2+no2tmean2, data = train_data)
summary(m1_pm10)

test_data<-chicago_table %>%
  filter(is.na(pm10tmean2))
test_data$pm10tmean2 <- predict(m1_pm10,newdata = test_data)

chicago_table_pm10 <- rbind(chicago_table,test_data)

#model to predict pm25tmean2
m2_pm25 <- lm(pm25tmean2 ~ tmpd+dptp+o3tmean2+no2tmean2+pm10tmean2,data = train_data)
summary(m2_pm25)

test_data_1<-chicago_table_pm10 %>%
  filter(is.na(pm25tmean2))

test_data_1$pm25tmean2 <- predict(m2_pm25,newdata = test_data_1)

chicago_clean <- na.omit(rbind(chicago_table,test_data,test_data_1))

# Outlier Treatment - capping
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 3 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- qnt[1] - H
  y[x > (qnt[2] + H)] <- qnt[2] + H
  y
}

chicago_clean$tmpd <- remove_outliers(chicago_clean$tmpd)
chicago_clean$dptp <- remove_outliers(chicago_clean$dptp)
chicago_clean$pm25tmean2 <- remove_outliers(chicago_clean$pm25tmean2)
chicago_clean$pm10tmean2 <- remove_outliers(chicago_clean$pm10tmean2)
chicago_clean$o3tmean2 <- remove_outliers(chicago_clean$o3tmean2)
chicago_clean$no2tmean2 <- remove_outliers(chicago_clean$no2tmean2)
```

3. Extract and display those rows (dplyr will display first 10 rows and that is fine) that have a 'seemingly' high value of PM2.5. Note, I am deliberately using managerial speak here when I say 'seemingly' but its your job to translate this into something convincing. Think about dsiplaying 10 rows that have most meaning.

```{r,warning=FALSE}
ggplot(chicago_clean,aes(pm25tmean2))+geom_histogram()
```

We can see that pm25tmean follows a normal distribution. So ~99.7% values lie between mean+/-3*sigma. So, pm25tmean2 values which are more than 3 standard deviations away from the mean are considered as high values

```{r}
cutoff <- mean(chicago_clean$pm25tmean2) + 3*sd(chicago_clean$pm25tmean2)

high_pm25 <- chicago_clean %>%
                   filter(pm25tmean2>cutoff)

high_pm25

```

4. Explore this subset of high PM2.5 level days further and see if other factors in the data could be linked to these high PM2.5 days. What queries would you run?

```{r}
ggplot(high_pm25, aes(pm10tmean2,pm25tmean2))+geom_line(size=1) 
```

```{r}
ggplot(high_pm25, aes(tmpd,pm25tmean2))+geom_line(size=1) 
```

```{r}
ggplot(high_pm25, aes(dptp,pm25tmean2))+geom_line(size=1) 
```

```{r}
ggplot(high_pm25, aes(o3tmean2,pm25tmean2))+geom_line(size=1) 
```

```{r}
ggplot(high_pm25, aes(no2tmean2,pm25tmean2))+geom_line(size=1) 
```
No link between high pm25 and other factors exists

##Correlation comparison
```{r}
cor(high_pm25[sapply(high_pm25, is.numeric)],use = "pairwise.complete.obs")
```

```{r}
cor(chicago_table[sapply(chicago_table, is.numeric)],use = "pairwise.complete.obs")
```
In the original data, pm25 is ~60% correlated with pm25 while in high pm25 value data the correlation is only 18%.Even in the case of other factors,the amount of correlation seems to be less compared to that of the original data.

5. Create two new columns that normalize the PM2.5 column using z-score and the PM10 column using min-max normalization

```{r}
chicago_clean <- chicago_clean %>%
  mutate(pm25_norm = (pm25tmean2-mean(pm25tmean2))/sd(pm25tmean2),
         pm10_norm = (pm10tmean2-min(pm10tmean2))/(max(pm10tmean2)
                                                   -min(pm10tmean2)))
```

6.  Using the command`year = as.POSIXlt(chicago$date)$year + 1900` add a new column for the year of the observation.

```{r}
chicago_clean$year = as.POSIXlt(chicago_clean$date)$year + 1900
```

7. Present key summary statistics of the pollutants by year 

```{r}
pm25_stats <- chicago_clean %>%
  group_by(year) %>% 
  summarise(avg = mean(pm25tmean2),median(pm25tmean2),max(pm25tmean2),min(pm25tmean2),sd(pm25tmean2))

pm10_stats <- chicago_clean %>%
  group_by(year) %>% 
  summarise(avg = mean(pm10tmean2),median(pm10tmean2),max(pm10tmean2),
            min(pm10tmean2),sd(pm10tmean2))

o3_stats <- chicago_clean %>%
  group_by(year) %>% 
  summarise(avg = mean(o3tmean2),median(o3tmean2),max(o3tmean2),min(o3tmean2),
            sd(o3tmean2))

no2_stats <- chicago_clean %>%
  group_by(year) %>% 
  summarise(avg = mean(no2tmean2),median(no2tmean2),max(no2tmean2),min(no2tmean2),sd(no2tmean2))

ggplot(pm25_stats,aes(year,avg)) +
geom_line(aes(color = "pm25")) +
geom_line(data = pm10_stats,aes(color = "pm10")) +
geom_line(data = o3_stats,aes(color = "o3")) +
geom_line(data = no2_stats,aes(color = "no2")) +
labs(title="Average pollutants by year", x="Year", color="Legend text")

```

1. The average of pollutants across years keeps decreasing for all of them from 1995  except o3. overall o3 shows an increase
 2. The yearly average of the pollutant no2 and pm10 follows a similar trend

8. What are the average levels of Ozone (o3) within quintiles of PM25? Note you have to convert pm25 levels into quintile groups and strore these levels into a new variable -- call it `pm25.quint`. Use the `cut` fucntion that divides the range of x into intervals and codes the values in x according to which interval they fall. 


```{r}
chicago_clean$pm25.quint <- with(chicago_clean,cut(pm25tmean2,breaks = quantile(
                          pm25tmean2, probs = seq(0,1,by=0.20),na.rm = TRUE),labels = c("1","2","3","4","5"),include.lowest = TRUE))

chicago_clean %>%
  group_by(pm25.quint) %>%
  summarise(mean(o3tmean2))

```
## The average o3 for the lowest 20% of pm10 values is high compare to the next 20% and so on. This trend is followed for 4 quintiles